---
title: "Week 7 Lab Session"
author: "Sung Lien A16628474"
date: "2025-02-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 1. Proportion og G/G in a population

Downloaded a CSV file from Ensemble < https://useast.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=959672880#373531_tablePanel

Here we read this CSV file 

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```


```{r}
table(mxl$Genotype..forward.strand.)
```


```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

Now let's look at a different population. I picked the GBR.

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```


Find proportion of G|G
```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) *100, 2)
```

This varient that is associated with childhood asthma is more frequent in the GBR population than the MKL population.

Lets now dig into this further.

## Section 4: Population Scale Analysis
One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale. 

How many samples do we have?

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```


```{r}
nrow(expr)
```


```{r}
table(expr$geno)
```

```{r}
library(ggplot2)
```

Lets make a boxplot

```{r}
ggplot(expr) + aes(geno, exp, fill=geno) +
  geom_boxplot(notch=TRUE)
```












