---
title: "Halloween Project Class 09"
author: "Sung Lien A16628474"
format: html
toc: true
---

Today we will examine data from 538 on common Hallloween candy. In particular we will use ggplot, dplyr, and PCA to make some of this multivariate dataset.

## Importing candy data

```{r}
candy <- read.csv("candy-data.csv", row.names=1)
head(candy)
```



>Q1 How many different candy types are in this data set?

```{r}
nrow(candy)
```


>Q2 How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

> How many chocolate candy are there in the dataset?

```{r}
sum(candy$chocolate)
```


>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Twix","winpercent"]
```

```{r}
candy["Twix",]$winpercent
```


```{r}
candy["Nerds",]$winpercent
```

```{r}
candy["3 Musketeers",]$winpercent
```

>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat",]$winpercent
```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars","winpercent"]
```

To get a quick overview of a new dataset the `skim` package can be useful:
```{r}
library("skimr")
skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

**N.B.** It looks like the `winpercent` column is on a different scale than the others (0-100% rather than 0-1). I will need to scale this dataset before analysis like PCA

>Q7. What do you think a zero and one represent for the candy$chocolate column?
They represent true or false whether the candy bar is a chocolate bar or not.

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

```{r}
library(ggplot2)


ggplot(candy) +
  aes(winpercent) +
  geom_histogram(bins=10, fill="lightblue")
```

>Q9. Is the distribution of winpercent values symmetrical?

No

>Q10. Is the center of the distribution above or below 50%?

Below

```{r}
summary(candy$winpercent)
```


>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

- Step 1: find all "chocolate" candy
- Step 2: find their "winpercent" values
- Step 3: summarize these values

- step 4: find all "fruity" candy
- step 5: find their winpercent values
- step 6: summarize these values

- step 7: compare the two summary values


1. Find all chocolate candy
```{r}
choc.inds <- candy$chocolate == 1
```

2. find their "winpercent" values
```{r}
choc.win <- candy[choc.inds,]$winpercent
```

3. summarize these values
```{r}
choc.mean <- mean(choc.win)
choc.mean
```

Do the same thing for fruit candy
```{r}
fruit.inds <- candy$fruity == 1
fruit.win <- candy[fruit.inds, "winpercent"]
fruit.mean <- mean(fruit.win)
mean(fruit.win)
```


Clearly chocolate has a higher mean `winpercent` than fruity candy
```{r}
choc.mean
fruit.mean
```

>Q12. Is this difference statistically significant?

```{r}
t.test(choc.win, fruit.win)
```
## Overall Candy Rankings


>Q13. What are the five least liked candy types in this set?

```{r}
# Not that useful -it just sorts the values
sort(candy$winpercent)
```

```{r}
x <- c(10, 1, 100)
order(x)
```

```{r}
x[ order(x)]
```

The `order()` function tells us how it arranges the elements of the input to make them sorted -i.e how to order them

we can determine the order of winpercent to make them sorted and use that order to arrange the whole dataset.

```{r}
ord.inds <- order(candy$winpercent)
head(candy[ord.inds, ])
```


> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[ord.inds, ])
```

```{r}
ord.inds <- order(candy$winpercent, decreasing = T)
head(candy[ord.inds, ])
```




>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}

ggplot(candy) +
  aes(winpercent, rownames(candy))+
  geom_col()
```


Let's rearrange 
```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent))+
  geom_col()
```

### Time to add some useful color


```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent), fill=chocolate)+
  geom_col()
```

We need to make our own separate color vector where we can spell out what candy is colored in a particular color.

```{r}
mycols <- rep("gray", nrow(candy))
mycols[candy$chocolate == 1] <- "chocolate"
mycols[candy$bar == 1] <- "brown"
mycols[candy$fruity == 1] <- "blue"
mycols
```

```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent))+
  geom_col(fill=mycols)
```


```{r}
as.logical(c(1,0,1))
```

Now, for the first time, using this plot we can answer questions like:


>Q17. What is the worst ranked chocolate candy?

Charleston Chew

>Q18. What is the best ranked fruity candy?

Nick L Lip

## Taking a look at pricepercent

Make a plot of winpercent (x-axis) vs pricepercent (y-axis)

```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=mycols) +
  geom_text()
```

To avoid the overplotting of the text labels we can use the add on package **ggrepel**

```{r}
library("ggrepel")
```

```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=mycols) +
  geom_text_repel(max.overlaps = 6) +
  theme_bw()
```


>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

```{r}
ord <- order(candy$pricepercent, decreasing = FALSE)
head( candy[ord,c(11,12)], n=5 )
```
Tootsie Roll Midgies

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
Nick L Lip



## 5 Exploring the correlation structure

Now that we have explored the dataset a little, we will see how the variables interact with one another.

First we will use correlation and view the results with the **corrplot** package to plot a correlation matrix.

```{r}
cij <- cor(candy)
cij
```



```{r}
library(corrplot)

corrplot(cij)
```


> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Chocolate and Fruity are anti-correlated

> Q23. Similarly, what two variables are most positively correlated?

Chocolate tends to be bars, have a high pricepercent and high winpercent.



## 6. Principal Component Analysis


Let’s apply PCA using the `prcom()` function to our candy dataset remembering to set the **scale=TRUE** argument.


```{r}
pca <- prcomp(candy, scale=TRUE)
```

```{r}
summary(pca)
```

```{r}
attributes(pca)
```


Let's plot our main results as our PCA "score plot"

```{r}
ggplot(pca$x) +
  aes(PC1, PC2, label=rownames(pca$x)) +
  geom_point(col=mycols)+
  geom_text_repel(col=mycols)
```




Finally let's look at how the original varaibles contribute to the PCs, start with PC1

```{r}
ggplot(pca$rotation) +
  aes(PC1, rownames(pca$rotation)) +
  geom_col()
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, pluribus, and hard, yes.