---
title: "Lab5_Data_Viz_with_ggplot"
author: "Sung Lien A16628474"
format: pdf
---
## Intro to ggplot

#

There are many graphics systems in R (ways to make plots and figures). These include "base" R plots. Today we will focus mostly on  **ggplot2** package.

Let's start with a plot of a simple in-built dataset called `cars`. 

```{r}
plot(cars)

```


Let's see how we make this figure using **ggplot**. First I need to install this package on my computer. To install any R package I use the function `install.packages()`. 

>I will run `install.packages("ggplots2") in my R console not this quarto document.



Before I can use any functions from add on packages I need to load package from my "library()" with the `library(ggplot2)` call. 

```{r}
library(ggplot2)
ggplot(cars)
```

All ggplot figures have at least 3 things (called layers). These include: 

-**data** (the input dataset I want to plot from), 
-**aes** (the aesthetic mapping of the data to my plot),
-**geoms** (the geom_point(), geom_line()).

```{r}
ggplot(cars) + aes(x=speed, y=dist) + geom_point()
```



Let's add a line to show the relationship here: 

```{r}
ggplot(cars) + aes(x=speed, y=dist) + geom_point() + geom_smooth(method = "lm") + theme_bw()+ labs(title = "My first GGPLOT")

```


>Q For which phases is data visualization important in our scientific workflows?
communication of results, exploratory data analysis. detection of outliers, all of the above
Answer: all of the above


>Q True or False? The ggplot2 package comes already installed with R? 
Answer: False

>Q Which plot types are typically NOT used to compare distributions of numeric variables? density plots, network graphs, histograms, violin plots, box plots
Answer: network graphs

>Q Which statement about data visualization with ggplot2 is incorrect?
Answer:ggplot is the only way to create plots in R

>Q Which gemotetic layer should be used to create scatter plots 

>geom_point

```{r} 
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```


> How many genes are in this dataset? 

```{r}
nrow(genes)
```

A first plot of this 

```{r}
ggplot(genes) + aes(x=Condition1, y=Condition2, col=State)+ geom_point() + theme_bw() + labs(title = "Gene Expression changes upon drug treatment", x="Control Expression", y="Drug Pressenece Expression") + scale_colour_manual(values=c("blue", "gray", "red"))
```




```{r}
table(genes$State)
```

> Q.Using your values above and 2 significant figures. What fraction of total genes is up-regulated in this dataset?

  
```{r}
round(table(genes$State)/nrow(genes), 4)
```

```{r}
n.tot <- nrow(genes)

vals <- table(genes$State)

vals.percent <- vals/n.tot * 100
round(vals.percent, 2) 
```


```{r}
colnames(genes)
```
```{r}
round(table(genes$State)/nrow(genes), 2)
```



```{r}
p <- ggplot(genes) + aes(x=Condition1, y=Condition2, col=State)+ geom_point()
```


```{r}
p + theme_minimal()
```


```{r}
p + labs(title = "Hello")
```


```{r}
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```

```{r}
library(dplyr)
library(gapminder)
```

```{r}
gapminder_2007 <-gapminder %>% filter(year==2007)
```

```{r}
ggplot(gapminder_2007) + aes(x=gdpPercap, y=lifeExp) + geom_point(alpha=0.5) + labs(title= "lifeExp/gdgPercap")

```

```{r}
gapminder_1957<-gapminder %>% filter(year==1957)
```

```{r}
ggplot(gapminder_1957) + 
  aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 7)
```

```{r}
ggplot(gapminder_2007) + aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) + geom_point(alpha=0.2) + labs(title= "lifeExp/gdpPercap") + scale_size_area(max_size = 7) + facet_wrap(~year)

```

```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)
```

```{r}
ggplot(gapminder_1957) + 
  geom_point(aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop), alpha=0.7) + 
  scale_size_area(max_size = 8) +
  facet_wrap(~year)
```


