---
title: "Comparative Structure Analysis"
author: "Sung Lien A16628474"
format: html
---

Here we run through a complete "pipline" of structure analysis that begins with a singe sequence identifier and ends with a PCA analysis.

```{r}
library(bio3d)
```

## Step 1. Get a sequence. Retrieve a sequence for the protein we are interested in. We will take ADK "1ake_A"

```{r}
id <- "1ake_A"
aa <- get.seq(id)
aa
```


## Step 2 Blast Search

Run a Blast search of the PDB for all related sequences to our input `aa`
```{r}
blast <- blast.pdb(aa)
```

```{r}
hits <- plot(blast)
```

## Step 3. Download all structures

These are our "top hits" i.e. all the structures in the PDB database related to our input sequence.
```{r}
hits$pdb.id
```


```{r}
# Download related PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```
## Step 4 Align and superpose

```{r}
pdbs <- pdbaln(files, fit=TRUE, exefile="msa")
```

```{r}
pdbs
```


## Step 5 PCA

Let's use our old friend PCA to make sense of these confusing, complicated structure relationships.

```{r}
pc <- pca(pdbs)
plot(pc)
```


Let's make a trajectory (or movie) of the main conformational changes captured by PC1. We will use the `mktrj()` function for this...

```{r}
mktrj(pc, file="pca_results.pdb")
```


Back of the envelope comparasion of the PDB size to UniProt

```{r}
uniprot <- 253206171 
pdb <- 231029
pdb/uniprot *100
```
